Stochastic Convolutional Sparse Coding \\ Supplementary Material

### 2019-08-19 VMV19/Results.tex
add directory

\section{Results} \label{sec:result}
We first validate the proposed algorithms on the fruit and city
datasets~\cite{zeiler2010deconvolutional}, which each consist of 10
training images of size $100 \times 100$. The online-mode algorithms
are then adapted to one thousand $100 \times 100$ image patches
randomly picked up from ImageNet~\cite{deng2009imagenet}. Note that
batch-based CSC commonly can only handle less than 100 images
simultaneously. The dictionary size is set to $100$ filters of size
$11 \times 11$ pixels in all experiments except for over-complete
dictionary. All training and evaluation processes in this manuscript
are performed on contrast normalized
images~\cite{zeiler2010deconvolutional,heide2015fast}.

\begin{figure}[h]
\centering
\begin{subfigure}{0.5\textwidth}
  \includegraphics[width=1\linewidth]{figure/iteVSobj.pdf}
\end{subfigure}

\begin{subfigure}{0.5\textwidth}
  \includegraphics[width=1\linewidth]{figure/timeVSobj.pdf}
  \vspace*{1mm}
\end{subfigure}

\begin{subfigure}{0.23\textwidth}
  \includegraphics[width=1\linewidth]{figure/batchFruit100.pdf}
\end{subfigure}
\begin{subfigure}{0.23\textwidth}
  \includegraphics[width=1\linewidth]{figure/heideFruit100.pdf}
\end{subfigure}
\caption{Top: Convergence comparison between the-state-of-art method~\cite{heide2015fast} and the proposed method with different subsampling rate, all of which on performed on fruit dataset. Convergence is evaluated by monitoring the objective value of Eq.~\ref{eq:CSCmodel} on training images versus iterations and time, respectively. Bottom: Learned filters by the proposed method with $p=0.1$ and the comparable method. In these represented learned filters, our method learns more smooth Gabor-like filters and less number of noise-like filters with the same $\lambda$. Beyond this, it runs faster with regard to each iteration and shows better convergence behavior.}
\label{fig:subsampleResult}
\end{figure}

\begin{figure*}[h]
    \centering
    \includegraphics[width=1\textwidth]{figure/reconImage.pdf}
    \resizebox{0.9\linewidth}{!}{
        \begin{tabular}{|c||c|c|c|c|c|c|c|c|c|c|}
            \cline{1-11}
            Image & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 \\
            \hline
            PSNR~\cite{heide2015fast} & 29.54 & 28.16 & 29.40 & 29.22 & 28.89 & 29.29 & 28.10  & 29.57 & 27.46 & 30.72 \\
            \hline
            PSNR ours & \textbf{29.65} & \textbf{28.18} & \textbf{29.52} & \textbf{29.43} & \textbf{28.99} & \textbf{29.38} & \textbf{28.19} & \textbf{29.63} & \textbf{27.62} & \textbf{30.98} \\
            \hline
        \end{tabular} }
    \caption{Numerical comparisons of the reconstruction quality obtained from the presented filters and its comparison. The reconstructions are performed on $50\%$ randomly observed images, with $\lambda = 0.4$ and $50$ ADMM iterations for both cases. Obtained PSNR values are averaged on 5 trials. Note that none of the testing images are in the training sets.} \label{fig:PSNRrecon}
\end{figure*}

@@

{\bfseries Convergence.} Comparisons of the convergence between the
proposed method (SBCSC) and the state-of-the-art batch-mode
algorithm~\cite{heide2015fast} are shown in
Fig.~\ref{fig:subsampleResult}. A different selection of the
subsampling rate reveals that the proposed strategy will slightly
influence the convergence and the training objective of the
minimization problem. Specifically, the more subsampled, the
relatively slower convergence and the higher objective will be
obtained. On the other hand, small subsampling rate will significantly
accelerate the computation process, where $10\%$ subsampling achieves
about $6 \times$ speedup over the not subsampled spatial domain solver
and a $2 \times$ speedup over state-of-the-art Fourier-domain solver
for one iteration. We observe that a subsampling rate between $p=0.1$
and $p=0.2$ delivers empirically good enough convergence in our
settings, as well as achieving at least $3 \times$ speedup. In
general, the proposed method with various subsampling rates converges
at around 10-12 iterations in all testing cases, acting similar to the
competing methods.

We have observed that the comparison method has a relatively
high objective value during the first $6-8$ iterations. The reason is that it introduces an additional mask term handling the boundary issues, therefore different splitting strategies are utilized and the subproblems are
interleaved on their primary variables (in some sense, the primary variables can be regarded as intermediate results in the ADMM iteration). See~\cite{wohlberg2016efficient} for details. Despite of this, it converges to an optimum with comparable objective after 10 iterations. In summary, the convergence behaviors of the proposed
algorithm is only slightly influenced by the subsampling strategy
within the testing subsampling rates. Comparing to the
state-of-the-art frequency solver, the proposed stochastic
spatial-domain solver with a subsampling rate of $0.1$ reduces the
computation time by a factor of two for the tested example. The
robustness of the proposed algorithm is evaluated by additional
experiments. Please see supplementary materials for reference.

{\bfseries Reconstruction}. The SBCSC-learned filters with a
subsampling rate of $p=0.1$ are shown in the bottom of
Fig.~\ref{fig:subsampleResult}. For a visual comparison, we also show
the filters learned from the competing method (bottom left of Figure 3
in~\cite{heide2015fast}). As can be observed, both of them learn some
seemingly similar Gabor-like filters. However, zooming into the
details (please see supplemental materials for high resolution
versions) reveals that our filters appear less noisy than the ones
from~\cite{heide2015fast}. We then demonstrate the effectiveness of
the reconstructed filters in the application of image inpainting,
which refers to reconstructing a full image from partial
measurements. A numerical comparison of the reconstruction quality is
shown in Fig.~\ref{fig:PSNRrecon}. The filters learned by the proposed
method not only appear less noisy by themselves, but also achieve
better reconstruction quality on partially observed images with
respect to the PSNR value. Furthermore, the learning process
only takes about $50\%$ of the time for our method. Specifically, SBCSC takes 170 seconds and the comparable method takes 350 seconds for 14 iterations on a Core i7 PC.

@@

{\bfseries Convergence.} Unlike the batch-based learning approaches
which evaluate its convergence by monitoring the objective value on
training datasets, a common way to evaluate the learning process of
online learning model is to monitor its objective value on test
datasets. In Fig.~\ref{fig:onlineSmall} we plot the objective values
against the iteration number for the proposed method (SOCSC) and a
recent online frequency-domain CSC method~\cite{liu-2018-first}. In
the same figure, we also keep track of the capability of the updated
filters during the learning process to sparsely represent the test
images, which is demonstrated by the time evolution of PSNR. These two
approaches stop at optimum positions with similar objective values.
%although they exhibit slightly different convergence behaviors with
%respect to iterations.
The final PSNR values for both methods also reveal a similar
reconstruction performance of the learned filters. In terms of runtime
comparison, however, the proposed method runs at least $6$ times faster
than the comparison method. Specifically, SOCSC takes 70 seconds and the comparable method takes 440 seconds on a Core i7 PC to process all 10 training images. The supplement shows additional comparisons for the other datasets.

@@

{\bfseries Over-complete dictionary.} To our knowledge, none of the
existing CSC work reported or analyzed the over-complete dictionary
(number of the dictionary is more than its degrees of freedom). One of
the reasons could be that most of the prior work is batch-based, thus
learning over-complete dictionary from small datasets would cause
overfitting issues, which may contain quite a few data-specific
filters, and therefore limit the ability to generalize the filters to
other data (we verify this explanation in the supplement). The
proposed online-based learning strategy (SOCSC) can overcome this
issue by scaling the model up to arbitrary sample sizes.

\begin{figure}[h]
\centering
\begin{subfigure}{0.49\textwidth}
  \includegraphics[width=1\linewidth]{figure/onlineVSliu-ite-fruit.pdf}
\end{subfigure}
\begin{subfigure}{0.49\textwidth}
  \includegraphics[width=1\linewidth]{figure/onlineVSliu-time-fruit.pdf}
\end{subfigure}

\caption{The experiments are performed on fruit dataset, and each iteration randomly choose one samples from the training datasets. Top: Convergence of the test set objectives (objective value of Eq.~\ref{eq:CSCmodel} on testing datasets) for our method (SOCSC) and the state-of-the-art online approach~\cite{liu-2018-first}. Bottom: Testing PSNR with respect to execution time. While the quality of the output is comparable, our method achieves $6 \times$ speedup.}
\label{fig:onlineSmall}
\end{figure}

@@

\begin{subfigure}{1\textwidth}
    \centering
  \includegraphics[width=0.75\linewidth]{figure/online100-filter.pdf}
\end{subfigure}
\end{minipage}
\begin{minipage}{0.6\textwidth}
\centering
\includegraphics[width=0.97\linewidth]{figure/online400-filter.pdf}
\vspace*{2mm}
\end{minipage}
\begin{minipage}{1\textwidth}
    \centering
    \resizebox{0.8\linewidth}{!}{
        \begin{tabular}{|c||c|c|c|c|c|c|c|c|c|c|}
            \cline{1-11}
            Image & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 \\
            \hline
            PSNR by (a) & 29.58 & 28.19 & 29.44 & 29.63 & 28.89 & 29.33 & 28.13  & 30.14 & 27.42 & 30.89 \\
            \hline
            PSNR by (b) & 29.63 & 28.22 & 29.57 & 29.90 & 29.12 & 29.59 & 28.05  & 30.17 & 27.53 & 31.08 \\
            \hline
            PSNR by (c) & \textbf{30.24} & \textbf{28.34} & \textbf{29.95}  & \textbf{30.30} & \textbf{29.43} & \textbf{29.96} & \textbf{28.24} & \textbf{30.57} & \textbf{27.72} & \textbf{31.67} \\
            \hline
        \end{tabular} }
\end{minipage}
\caption{Filters learned from 1000 images by our method (SOCSC) and the state-of-the-art online method~\cite{liu-2018-first}, and their corresponding reconstruction quality using the same hyperparameter settings as in Fig.~\ref{fig:PSNRrecon}. (a) The under-complete dictionary ($11 \times 11 \times 100$) learned by~\cite{liu-2018-first}; (b) The under-complete dictionary ($11 \times 11 \times 100$) learned by SOCSC. (c) The over-complete dictionary ($11 \times 11 \times 400$) learned by SOCSC. These under-complete dictionaries, mainly composed of Gabor-like filters, can be seen as a subset of the represented over-complete dictionary, which contains a number of extra low contrast image features.}
\label{fig:overCompleteDic}
\end{figure*}

We demonstrate this ability on 1000 image patches with the size of
$100 \times 100$, and learn an $11 \times 11 \times 400$ over-complete
dictionary, which is shown in Fig.~\ref{fig:overCompleteDic}. For a
visual comparison, we also show $100$ learned filters generated by the
same algorithm and another 100 filters generated
by~\cite{liu-2018-first}. As can be observed, both of the approaches
learn visually similar under-complete dictionary, while the proposed
method takes $6 \times$ less training time than the comparison
method. The learned over-complete dictionary is composed of
the Gabor-like image features as represented in the under-complete
dictionaries, as well as a number of low contrast features which are
not typical for under-complete dictionaries. This additional
feature information would play an essential role to better reconstruct
the natural images, which can be demonstrated by the results of
reconstruction quality shown at the bottom of the same figure.

We further show a bottleneck revealed by the under-complete dictionary
in the online-based CSC model. The top of
Fig.~\ref{fig:overComDicAndMinibatch} demonstrates that no more
apparent progress could be observed when the number of training images
is higher than a specific value for both of the online approaches
($K=100$). However, owing to more abundant filters, learning
over-complete dictionary overcomes this bottleneck, and it shows a
considerable improvement in the PSNR of image representations.  This
is also verified by applying these learned dictionaries to the image
inpainting application as discussed above.

@@

\begin{figure}[h]
\centering
\begin{subfigure}{0.5\textwidth}
  \includegraphics[width=1\linewidth]{figure/overComplete-ite.pdf}
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
  \includegraphics[width=1\linewidth]{figure/minibatch.pdf}
\end{subfigure}

@@

The runtime comparisons for various mini-batch sizes are shown in the
bottom of Fig.~\ref{fig:overComDicAndMinibatch}. Note that larger
$\eta$ will result in a smaller number of iterations to process all
$1000$ samples. The plots show that a larger mini-batch size will
generally lead to a greater progress in first few training steps,
though it takes additional running time and memory. Overall,
mini-batched updates provides a more runtime efficient learning
process in the online-based CSC model, and according to the obtained
experimental results, $\eta=20$ achieves a one order of magnitude
speedup over $\eta=1$ to reach a a comparable level of
convergence. Since computing sparse codes is a data-independent
process, this mini-batched approach can be further accelerated in a
distributed-computing system.

>>>>

\section{Experiments and Results} \label{sec:result}

\changed{ \subsection{Experimental Design}
We first validate the proposed algorithms on the fruit and city
datasets~\cite{zeiler2010deconvolutional}, where each consists of 10
training images of size $100 \times 100$. The online-mode algorithms
are then adapted to one thousand $100 \times 100$ image patches
randomly picked up from ImageNet~\cite{deng2009imagenet}. Note that batch-based CSC commonly can only handle less than 100 images
simultaneously due to the memory limitation. The dictionary size is set to $100$ filters of size $11 \times 11$ pixels in all experiments except for over-complete dictionary. All training and evaluation processes in this manuscript are performed on contrast normalized
images~\cite{zeiler2010deconvolutional,heide2015fast}.}

\begin{figure}[h]
\centering
\begin{subfigure}{0.45\textwidth}
  \includegraphics[width=1\linewidth]{figure/iteVSobj.pdf}
\end{subfigure}

\begin{subfigure}{0.45\textwidth}
  \includegraphics[width=1\linewidth]{figure/timeVSobj.pdf}
  \vspace*{1mm}
\end{subfigure}

%\begin{subfigure}{0.23\textwidth}
%  \includegraphics[width=1\linewidth]{figure/batchFruit100.pdf}
%\end{subfigure}
%\begin{subfigure}{0.23\textwidth}
%  \includegraphics[width=1\linewidth]{figure/heideFruit100.pdf}
%\end{subfigure}
\caption{Convergence comparison between the-state-of-art method~\cite{heide2015fast} and the proposed method with different subsampling rate, all of which are performed on fruit dataset. Convergence is evaluated by monitoring the objective value of Eq.~\ref{eq:CSCmodel} on training images versus iterations and time, respectively.}
\label{fig:subsampleResult}
\end{figure}

@@

{\bfseries Convergence.} Comparisons of the convergence between the
proposed method (SBCSC) and the state-of-the-art batch-mode
algorithm~\cite{heide2015fast} are shown in
Fig.~\ref{fig:subsampleResult} \changed{(the compared method uses a similar number of iterations as ours to reach convergence)}. A different selection of the
subsampling rate reveals that the proposed strategy will slightly
influence the convergence and the training objective of the
minimization problem. Specifically, the more subsampled, the
relatively slower convergence and the higher objective will be
obtained. On the other hand, small subsampling rate will significantly
accelerate the computation process, where $10\%$ subsampling achieves
about $6 \times$ speedup over the not subsampled spatial domain solver
and a $2 \times$ speedup over state-of-the-art Frequency-domain solver
for one iteration. We observe that a subsampling rate between $p=0.1$
and $p=0.2$ delivers empirically good enough convergence in our
settings, as well as achieving at least $3 \times$ speedup. In
general, the proposed method with various subsampling rates converges
at around 10-12 iterations in all testing cases, acting similar to the
competing methods.

%We have observed that the comparison method has a relatively
%high objective value during the first $6-8$ iterations. The reason is that it introduces an additional mask term handling the boundary issues, therefore different splitting strategies are utilized and the subproblems are
%interleaved on their primary variables (in some sense, the primary variables can be regarded as intermediate results in the ADMM iteration). See~\cite{wohlberg2016efficient} for details. Despite of this, it converges to an optimum with comparable objective after 10 iterations.
In summary, the convergence behaviors of the proposed
algorithm is only slightly influenced by the subsampling strategy
within the testing subsampling rates. Comparing to the
state-of-the-art frequency solver, the proposed stochastic
spatial-domain solver with a subsampling rate of $0.1$ reduces the
computation time by a factor of two for the tested example. Specifically,
SBCSC takes 170 seconds and the comparable method takes 350 seconds for
14 iterations on a Core i7 PC. The
robustness of the proposed algorithm is evaluated by additional
experiments. Please see supplementary materials for reference.

%{\bfseries Reconstruction}. The SBCSC-learned filters with a
%subsampling rate of $p=0.1$ are shown in the bottom of
%Fig.~\ref{fig:subsampleResult}. For a visual comparison, we also show
%the filters learned from the competing method (bottom left of Figure 3
%in~\cite{heide2015fast}). As can be observed, both of them learn some
%seemingly similar Gabor-like filters. However, zooming into the
%details (please see supplemental materials for high resolution
%versions) reveals that our filters appear less noisy than the ones
%from~\cite{heide2015fast}. We then demonstrate the effectiveness of
%the reconstructed filters in the application of image inpainting,
%which refers to reconstructing a full image from partial
%measurements. A numerical comparison of the reconstruction quality is
%shown in Fig.~\ref{fig:PSNRrecon}. The filters learned by the proposed
%method not only appear less noisy by themselves, but also achieve
%better reconstruction quality on partially observed images with
%respect to the PSNR value. Furthermore, the learning process
%only takes about $50\%$ of the time for our method. Specifically, SBCSC takes 170 seconds and the comparable method takes 350 seconds for 14 iterations on a Core i7 PC.

@@

\begin{figure}[h]
\centering
\begin{subfigure}{0.45\textwidth}
  \includegraphics[width=1\linewidth]{figure/onlineVSliu-ite-fruit.pdf}
\end{subfigure}
\begin{subfigure}{0.45\textwidth}
  \includegraphics[width=1\linewidth]{figure/onlineVSliu-time-fruit.pdf}
\end{subfigure}

\caption{The experiments are performed on fruit dataset, and each iteration randomly choose one samples from the training datasets. Top: Convergence of the test set objectives (objective value of Eq.~\ref{eq:CSCmodel} on testing datasets) for our method (SOCSC) and the current online approach~\cite{liu-2018-first}. Bottom: Testing PSNR with respect to execution time. While the quality of the output is comparable, our method achieves $6 \times$ speedup.}
\label{fig:onlineSmall}
\end{figure}

{\bfseries Convergence.} Unlike the batch-based learning approaches
which evaluate its convergence by monitoring the objective value on
training datasets, a common way to evaluate the learning process of
online learning model is to monitor its objective value on test
datasets. In Fig.~\ref{fig:onlineSmall} we plot the objective values
against the iteration number for the proposed method (SOCSC) and a
recent online frequency-domain CSC method~\cite{liu-2018-first} (both
 approaches use Matlab built-in functions only). In
the same figure, we also keep track of the capability of the updated
filters during the learning process to sparsely represent the test
images, which is demonstrated by the time evolution of PSNR (PSNR is the 
peak signal-to-noise ratio, measuring the difference between the 
reconstructed signal and the original one). These two
approaches stop at optimum positions with similar objective values.
%although they exhibit slightly different convergence behaviors with
%respect to iterations.
The final PSNR values for both methods also reveal a similar
reconstruction performance of the learned filters. In terms of runtime
comparison, however, the proposed method runs at least $6$ times faster
than the comparison method. Specifically, SOCSC takes 70 seconds and the comparable method takes 440 seconds on a Core i7 PC to process all 10 training images. The supplement shows additional comparisons for the other datasets.

@@

\begin{figure}[h]
\centering
\begin{subfigure}{0.45\textwidth}
  \includegraphics[width=1\linewidth]{figure/nonZeroElement.pdf}
\end{subfigure}
\begin{subfigure}{0.45\textwidth}
  \includegraphics[width=1\linewidth]{figure/reconPSNR.pdf}
\end{subfigure}

\caption{Top: number of non-zero elements in the codes for different images (on average $0.2\%$ of the variables are non-zero when $K=100$). Bottom: PSNR between the reconstructed images and the original ones with under-complete and over-complete dictionaries, respectively.}
\label{fig:overVSunder}
\end{figure}

@@

\begin{subfigure}{1\textwidth}
    \centering
  \includegraphics[width=0.75\linewidth]{figure/online100-filter.pdf}
\end{subfigure}
\end{minipage}
\begin{minipage}{0.6\textwidth}
\centering
\includegraphics[width=0.97\linewidth]{figure/online400-filter.pdf}
\vspace*{2mm}
\end{minipage}
\begin{minipage}{1\textwidth}
\centering
\includegraphics[width=0.9\textwidth]{figure/reconImage.pdf}
\vspace*{2mm}
\end{minipage}
\begin{minipage}{1\textwidth}
    \centering
    \resizebox{0.8\linewidth}{!}{
        \begin{tabular}{|c||c|c|c|c|c|c|c|c|c|c|}
            \cline{1-11}
            Image & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 \\
            \hline
            PSNR by (a) & 29.58 & 28.19 & 29.44 & 29.63 & 28.89 & 29.33 & 28.13  & 30.14 & 27.42 & 30.89 \\
            \hline
            PSNR by (b) & 29.63 & 28.22 & 29.57 & 29.90 & 29.12 & 29.59 & 28.05  & 30.17 & 27.53 & 31.08 \\
            \hline
            PSNR by (c) & \textbf{30.24} & \textbf{28.34} & \textbf{29.95}  & \textbf{30.30} & \textbf{29.43} & \textbf{29.96} & \textbf{28.24} & \textbf{30.57} & \textbf{27.72} & \textbf{31.67} \\
            \hline
        \end{tabular} }
\end{minipage}
\caption{Top: Filters learned from large-scale datasets by our method (SOCSC) and the comparable online method~\cite{liu-2018-first}. Bottom: $10 ~ 256 \times 256$ testing images and their corresponding reconstruction quality in the image inpainting application. (a) The under-complete dictionary ($11 \times 11 \times 100$) learned by~\cite{liu-2018-first}; (b) The under-complete dictionary ($11 \times 11 \times 100$) learned by SOCSC. (c) The over-complete dictionary ($11 \times 11 \times 400$) learned by SOCSC. These under-complete dictionaries, mainly composed of Gabor-like filters, can be seen as a subset of the represented over-complete dictionary, which contains a number of extra low contrast image features.}
\label{fig:overCompleteDic}
\end{figure*}

{\bfseries Over-complete dictionary.}
%To our knowledge, none of the
%existing CSC work reported or analyzed the over-complete dictionary
%(number of the dictionary is more than its degrees of freedom). One of
%the reasons could be that most of the prior work is batch-based, thus
%learning over-complete dictionary from small datasets would cause
%overfitting issues, which may contain quite a few data-specific
%filters, and therefore limit the ability to generalize the filters to
%other data (we verify this explanation in the supplement). The
%proposed online-based learning strategy (SOCSC) can overcome this
%issue by scaling the model up to arbitrary sample sizes.
Learning over-complete dictionary (number of the dictionary is more
than its degrees of freedom) from small datasets would cause
overfitting issues, which may contain quite a few data-specific
filters, and therefore limit the ability to generalize the filters to
other data (we verify this explanation in the supplement). The
proposed online-based learning strategy (SOCSC) can overcome this
issue by scaling the model up to arbitrary sample sizes.

We demonstrate this ability on 1000 image patches with the size of
$100 \times 100$, and learn an $11 \times 11 \times 400$ over-complete
dictionary, which is shown in Fig.~\ref{fig:overCompleteDic}. For a
visual comparison, we also show $100$ learned filters generated by the
same algorithm and another 100 filters generated
by~\cite{liu-2018-first}. As can be observed, both of the approaches
learn visually similar under-complete dictionary, while the proposed
method takes $6 \times$ less training time than the comparison
method. The learned over-complete dictionary is composed of
the Gabor-like image features as represented in the under-complete
dictionaries, as well as a number of low contrast features which are
not typical for under-complete dictionaries. This additional
feature information would play an essential role to reveal an improved sparse representation of
the natural images. The numerical comparisons of number of non-zero elements and its corresponding reconstruction PSNR for testing images ($10 ~ 256 \times 256$ images presented in Fig.~\ref{fig:overCompleteDic}) are shown in Fig.~\ref{fig:overVSunder}. Here, we define the non-zero elements as the codes whose coefficient is no less than $0.1$. We could observe that at all times, using over-complete dictionary leads to a sparser representation of the images, roughly $8\%-10\%$ reduction on the non-zero elements. Meanwhile, it achieves dramatically improved reconstruction quality, over $1$ dB on average.

We further demonstrate the effectiveness of the over-complete
dictionary in the application of image inpainting,
which refers to reconstructing a full image from partial
measurements. A numerical comparison of the reconstruction quality is
shown in the bottom of Fig.~\ref{fig:overCompleteDic}. The reconstructions
are performed on $50\%$ randomly observed images, with $\lambda = 0.4$ and
$50$ ADMM iterations for all cases. Obtained PSNR values are averaged on
5 trials. The over-complete filters learned by the proposed
method achieves significantly improved reconstruction quality on partially 
observed images in terms of the PSNR value.

We also observe a bottleneck revealed by the under-complete dictionary
in the online-based CSC model. The top of Fig.~\ref{fig:overComDicAndMinibatch} demonstrates that no more
apparent progress could be observed when the number of training images
is higher than a specific value for both of the online approaches
($K=100$). However, owing to more abundant filters, learning
over-complete dictionary overcomes this bottleneck, and it shows a
considerable improvement in the PSNR of image representations. All 
presented experimental results imply that the number of filters and 
number of training samples are both essential in the CSC model.

@@

\begin{figure}[h]
\centering
\begin{subfigure}{0.45\textwidth}
  \includegraphics[width=1\linewidth]{figure/overComplete-ite.pdf}
\end{subfigure}
\begin{subfigure}{0.45\textwidth}
  \includegraphics[width=1\linewidth]{figure/minibatch.pdf}
\end{subfigure}

@@

The runtime comparisons for various mini-batch sizes are shown in the
bottom of Fig.~\ref{fig:overComDicAndMinibatch}. Note that larger
$\eta$ will result in a smaller number of iterations to process all
$1000$ samples. The plots show that a larger mini-batch size will
generally lead to a greater progress in first few training steps,
though it takes additional running time and memory. Overall,
mini-batched update provides a more runtime efficient learning
process in the online-based CSC model, and according to the obtained
experimental results, $\eta=20$ achieves one order of magnitude
speedup over $\eta=1$ to reach a a comparable level of
convergence. Since computing sparse codes is a data-independent
process, this mini-batched approach can be further accelerated in a
distributed-computing system.

### 2019-08-19 VMV19/SCSCmodel.tex
add directory

\section{Stochastic Convolutional Sparse Coding}
\subsection{The Model}
We first define a stacked vector $\code = [\code_1, \dots, \code_K]
\in \mathbb{R}^{DK}$ for the codes, as well as a stacked matrix
$\Filter = [\Filter_1, \dots, \Filter_K] \in \mathbb{R}^{D \times DK}$
for the filter convolutions. The convolution operators are expressed
as a matrix multiplication so that $ \Filter \code = \sum_{k=1}^{K}
\filter_k * \code_k$.  Therefore, each part of $\Filter$ is a Toeplitz
matrix.

@@

After computing the subsampled codes $\subcode^\ite$, we can then project them
onto the original spatial support by
\begin{equation} \label{eq:upsampleCode}
    \code^\ite = (\mask^{\ite})^\top \subcode^\ite.
\end{equation}
Afterwards, the dictionary can be updated by solving the optimization problem
\begin{equation} \label{eq:updatingFilter}
\begin{split}
   & \filter^{\ite} = \minimize{\filter} \frac{1}{2}\|\signal - \Code^{\ite} \filter \|_2^2 \\
   & \text{subject to}  ~ \|\filter_k\|^2_2 \leq 1 ~ \forall k \in \{1,\dots,K\},
\end{split}
\end{equation}
where $\Code^{\ite}= [\Code_1^{\ite}, \dots, \Code_K^{\ite}]
\in \mathbb{R}^{D \times MK}$ is a concatenation of Toeplitz matrices,
and $\Code_k^{\ite}$ is constructed from the associated $\code_k^{\ite}$, $\filter=
[\filter_1,\dots,\filter_K] \in \mathbb{R}^{MK}$ such that $ \Code
\filter = \sum_{k=1}^{K} \filter_k * \code_k$. In typical CSC
settings, $M \ll D$, hence there is no need to perform subsampling on
the dictionary.

The work of~\cite{mensch2016dictionary} also applies a so-called
stochastic subsampling strategy for dictionary learning, where the
subsampling is randomly performed on the observed signals, differing
from that of our proposed. In addition, it is basically applied for
factorizing matrix that is large in both dimensions. Although CSC can
be expressed in a similar mathematical formula as matrix factorization, the two
problems are in practice quite different due to the convolutional
property of the CSC model. Our proposed randomization
approach shares a similar idea as utilized in the stochastic Frank-Wolfe
algorithm~\cite{reddi2016stochastic,pmlr-v80-kerdreux18a}, in which
the optimization is performed on a subset of the variables which are
randomly extracted based on a certain probability distribution at each
iteration. For the proposed algorithm, each iteration extracts
($pDK+MK$) variables, where $pDK$ variables are randomly picked from a
total $DK$ variables, and the rest remain unchanged. Owing to the fact
that the codes are highly sparse, the original signals can still be
represented by a portion of the codes under a reasonable subsampling
rate. Therefore, the convergence of the proposed algorithm will not be
significantly affected by the subsampling manipulation, unlike the
general case of the stochastic Frank-Wolfe algorithm, which usually
requires more iterations to reach convergence. This insight is
experimentally verified in Section~\ref{sec:result}.

@@

Problem~\eqref{eq:updatingCode} is the standard LASSO, which can be
solved by plenty of optimization frameworks. We found that solving it
with ADMM delivers a good balance between computation time and
convergence within a moderate number of iterations. Specifically, the
data fitting term and the $L_1$ penalty term are split, forming two
separate substeps. The first substep is a quadratic programming (QP)
problem, and we can either cache the matrix factorization by Cholesky
decomposition (when $N$ is relatively large), or solve it iteratively
by Conjugate Gradient (when $N$ is relatively small). The second
substep can be solved by a point-wise shrinkage operation.
%The above two substeps alternatively solved within the ADMM
%iterations.
Problem~\eqref{eq:updatingFilter} is a quadratic constrained
quadratic programming (QCQP) problem, and it can be efficiently solved
by projected block coordinate decent. Empirically, a single iteration
is enough with $\filter$ computed in previous iteration as a warm
start. We set the hyperparameters $\lambda=1$, the ADMM iteration
fixed to 10, the augmented Lagrangian penalty $\rho$ to $10 \lambda$,
and the over-relaxation strategy within ADMM is applied with $\alpha =
1.8$. For a detailed description of the above two solvers, please
refer to the supplement.

@@

\subsection{Complexity Analysis}
Recall that $D$ is the number of pixels for a single image, $K$ is the
number of filters, and $M$ is the size of the filter
support. Commonly, we can assume $K \approx M$.  State-of-the-art
frequency-domain solvers then have the time complexity
$\mathcal{O}(K^2D + KDlog(D))$ for a single data pass.

>>>>

\section{Stochastic Convolutional Sparse Coding}
\subsection{The SCSC Model}
We first define a stacked vector $\code = [\code_1, \dots, \code_K]
\in \mathbb{R}^{DK}$ for the codes, as well as a stacked matrix
$\Filter = [\Filter_1, \dots, \Filter_K] \in \mathbb{R}^{D \times DK}$
for the filter convolutions. The convolution operators are expressed
as a matrix multiplication so that $ \Filter \code = \sum_{k=1}^{K}
\filter_k * \code_k$.  Therefore, each part of $\Filter$ is a Toeplitz
matrix.

@@

After computing the subsampled codes $\subcode^\ite$, we can then project them
onto the original spatial support by
\begin{equation} \label{eq:upsampleCode}
    \code^\ite = (\mask^{\ite})^\top \subcode^\ite.
\end{equation}
Afterwards, the dictionary can be updated by solving the optimization problem
\begin{equation} \label{eq:updatingFilter}
\begin{split}
   & \filter^{\ite} = \minimize{\filter} \frac{1}{2}\|\signal - \Code^{\ite} \filter \|_2^2 \\
   & \text{subject to}  ~ \|\filter_k\|^2_2 \leq 1 ~ \forall k \in \{1,\dots,K\},
\end{split}
\end{equation}
where $\Code^{\ite}= [\Code_1^{\ite}, \dots, \Code_K^{\ite}]
\in \mathbb{R}^{D \times MK}$ is a concatenation of Toeplitz matrices,
and $\Code_k^{\ite}$ is constructed from the associated $\code_k^{\ite}$, $\filter=
[\filter_1,\dots,\filter_K] \in \mathbb{R}^{MK}$ such that $ \Code
\filter = \sum_{k=1}^{K} \filter_k * \code_k$. In typical CSC
settings, $M \ll D$, hence there is no need to perform subsampling on
the dictionary. Notice that the time complexity of solving the $\filter$-update 
step in spatial domain is dependent of $M$, in contrast with that of frequency-domain 
solvers, which is $D$ dependent. This addresses the third issue in Section~\ref{CSCmodel}.

%The work of~\cite{mensch2016dictionary} also applies a so-called
%stochastic subsampling strategy for dictionary learning, where the
%subsampling is randomly performed on the observed signals, differing
%from that of our proposed. In addition, it is basically applied for
%factorizing matrix that is large in both dimensions. Although CSC can
%be expressed in a similar mathematical formula as matrix factorization, the two
%problems are in practice quite different due to the convolutional
%property of the CSC model.
\changed{Our proposed randomization approach shares a similar idea as utilized in the stochastic Frank-Wolfe
algorithm~\cite{reddi2016stochastic,pmlr-v80-kerdreux18a}.  The general idea of that is to solve the optimization problem on a subset of the variables at each iteration, which are
randomly extracted based on a certain probability distribution.} For the proposed algorithm, each iteration extracts
($pDK+MK$) variables, where $pDK$ variables are randomly picked from a
total $DK$ variables, and the rest remain unchanged. Owing to the fact
that CSC model is \changed{over-parameterized} and the codes are highly 
sparse, the original signals can still be represented by a portion of 
the codes under a reasonable subsampling
rate. Therefore, the convergence of the proposed algorithm will not be
significantly affected by the subsampling manipulation, unlike the
general case of the stochastic Frank-Wolfe algorithm, which usually
requires more iterations to reach convergence. This insight is
experimentally verified in Section~\ref{sec:result}.

@@

Problem~\eqref{eq:updatingCode} is the standard LASSO, which can be
solved by plenty of optimization frameworks. We found that solving it
with ADMM delivers a good balance between computation time and
convergence within a moderate number of iterations. Specifically, the
data fitting term and the $L_1$ penalty term are split, forming two
separate substeps. The first substep is a quadratic programming (QP)
problem, and we can either cache the matrix factorization by Cholesky
decomposition (when $N$ is relatively large), or solve it iteratively
by Conjugate Gradient (when $N$ is relatively small). The second
substep can be solved by a point-wise shrinkage operation.
%The above two substeps alternatively solved within the ADMM
%iterations.
Problem~\eqref{eq:updatingFilter} is a quadratic constrained
quadratic programming (QCQP) problem, and it can be efficiently solved
by projected block coordinate decent. Empirically, a single iteration
is enough with $\filter$ computed in previous iteration as a warm
start. We set the \changed{hyper-parameters} $\lambda=1$, the ADMM iteration
fixed to 10, the augmented Lagrangian penalty $\rho$ to $10 \lambda$,
and the over-relaxation strategy within ADMM is applied with $\alpha =
1.8$. For a detailed description of the above two solvers, please
refer to the supplement.

@@

\subsection{Complexity Analysis}\label{complexity}
Recall that $D$ is the number of pixels for a single image, $K$ is the
number of filters, and $M$ is the size of the filter
support. Commonly, we can assume $K \approx M$.  State-of-the-art
frequency-domain solvers then have the time complexity
$\mathcal{O}(K^2D + KDlog(D))$ for a single data pass.
